Experiences with Planning for Natural Language Generation

### 2009-06-02 trunk/Papers/ci-spark/domains.tex
more changes to section 3

Let us now turn to a second recent application of planning in NLG, the
GIVE Challenge (``Generating Instructions in Virtual Environments'';
\citealt{ByrKolStrCasDalMooObe09}). The object of this shared task is
to build an NLG system which produces natural-language instructions
which will guide a human user in performing some task in a virtual
environment.  From an NLG perspective, GIVE makes for an interesting
challenge because it is a theory-neutral task that exercises all
components of an NLG system, and emphasizes the study of communication
in a (simulated) physical environment. Furthermore, because the client
which displays the 3D environment to the user and the NLG system can
be physically separated, as long as they are connected over a network,
such systems can be cheaply evaluated over the Internet. This provides
a potential solution to the long-standing problem of evaluating NLG
systems. The first installment of GIVE (GIVE-1) evaluated five NLG
systems on the performance of 1143 users, making it the largest NLG
evaluation effort in terms of human users ever.

\begin{figure}[t]
\centering
%\includegraphics[width=1 \columnwidth]{give_world_no_expl}
\includegraphics[width=0.75\columnwidth]{give_world_2}
\caption{Map of a GIVE world.}
  \label{fig:give-development-world}
\end{figure}

Planning plays a central role in the GIVE task. To see this, consider
the example GIVE world map in
Figure~\ref{fig:give-development-world}. In this example world, the
user's task is to pick up a trophy in the top left room. The trophy is
hidden in a safe behind a picture; to access it, the user must push
certain buttons in order to move the picture out of the way, open the
safe, and open doors. The user must to navigate the world and perform
these actions in the 3D client in a certain order; the NLG system must
instruct the user on how to do this.  To simplify both the planning
and the NLG task, the world is discretised into a set of tiles of
equal size. The user can turn by 90 degree steps in either direction,
and can move from the centre of one tile to the centre of the next
tile, provided the path between two tiles is not blocked.
Figure~\ref{fig:give-planning} shows the encoding of some of the
available GIVE domain actions, in PDDL syntax. In the example, the
shortest plan to solve the task consists of 108 action steps, with the
first few steps as follows:
%
\begin{enumerate}
\item $\mathsf{turn}\textsf{-}\mathsf{left}(\mathsf{north},
\mathsf{west})$,
\item $\mathsf{move}(\mathsf{pos\_5\_2}, \mathsf{pos\_4\_2}, \mathsf{west})$,
\item $\mathsf{manipulate}\textsf{-}\mathsf{b1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_5\_2})$,
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{west}, \mathsf{north})$.
\end{enumerate}

\begin{figure}
\centering
\begin{minipage}{0.5\textwidth}
{\small%
\begin{verbatim}
(:action move
   :parameters (?from - position
                ?to - position
                ?ori - orientation)
   :precondition 
       (and (player-pos ?from) 
            (adjacent ?from ?to ?ori) 
            (player-orient ?ori)
            (not-blocked ?from ?to)
            (not-alarmed ?to))
   :effect 
       (and (not (player-pos ?from))
            (player-pos ?to)))

(:action turn-left
   :parameters (?ori - orientation
                ?newOri - orientation)
   :precondition 
       (and (player-orient ?ori)
            (next-orient-left ?ori ?newOri))
   :effect 
       (and (not (player-orient ?ori))
            (player-orient ?newOri)))

(:action turn-right
   :parameters (?ori - orientation
                ?newOri - orientation)
   :precondition 
       (and (player-orient ?ori)
            (next-orient-right ?ori ?newOri))
   :effect 
       (and (not (player-orient ?ori))
            (player-orient ?newOri)))

(:action manipulate-b1-off-on
   :parameters (?pos - position)
   :precondition 
       (and (state b1 off)
            (player-pos ?pos)
            (position b1 ?pos))
   :effect 
       (and (not (state b1 off))
            (state b1 on)
            (not (state d1 closed))
            (state d1 open) 
            (not (blocked pos_6_5 pos_6_4))
            (not (blocked pos_6_4 pos_6_5))))
\end{verbatim}}%
\end{minipage}
\caption{Some PDDL actions for the GIVE domain.}
\label{fig:give-planning}
\end{figure}

A GIVE NLG system must be able to compute such plans. At a minimum,
the discourse planner will call a planner in order to determine the
content of the instructions that should be presented to the
user. Under this view, the planning problem is very similar to the
Gridworld problem (see, e.g., \citealt{Tovey-Koenig:2000}), which also
involves finding a route through a two-dimensional world map with
discrete positions. As in Gridworld, the domain requires not only
route finding in the discretised world, but also the execution of
certain actions (such as finding keys and unlocking doors in
Gridworld, or pushing the correct buttons to open doors and the safe
in GIVE). However, the GIVE worlds tend to be bigger than the
Gridworld instances that were used in the 1990s planning competitions.

This relatively loose integration of NLG system and planner is the
state of the art of the systems that participated in GIVE-1. However,
it is desirable to integrate the planner and the generation system
more closely than this. For instance, let's say that the NLG system
wants to generate the instruction sequence ``walk to the centre of the
room; turn right; now press the green button in front of
you''. Experiments with human instruction givers \cite{stoia:08} show
that this is a pattern that they use frequently: The instruction
follower is made to walk to a certain point in the world at which the
instruction giver can then use a a referring expression (``the green
button'') that is very easy to interpret. Such an NLG system must
integrate discourse planning and planning in the domain of the world
map closely. On the one hand, the structure of the discourse is
determined by the needs of the NLG system rather than the domain plan;
on the other hand, the discourse planner must be aware of the way in
which the instruction ``turn right'' is likely to change the
visibility of objects. Even if the NLG system doesn't implement the
generation of this discourse as planning, it still solves a problem
that subsumes the domain planning problem. For these reasons, we
consider the GIVE domain planning problem as a natural part of a GIVE
NLG system.  \todo{Make sure that this paragraph still talks about the
  things that the reviewers found interesting about it.}

>>>>

In the second application of planning in NLG, we consider the recent GIVE
Challenge (``Generating Instructions in Virtual Environments'';
\citealt{ByrKolStrCasDalMooObe09}). The object of this shared task is to
build an NLG system which produces natural language instructions which
guide a human user in performing a task in a virtual environment. From an
NLG perspective, GIVE makes for an interesting challenge since it is a
theory-neutral task that exercises all components of an NLG system, and
emphasizes the study of communication in a (simulated) physical
environment. Furthermore, because the client displaying the 3D environment
to the user can be physically separated from the NLG system (provided they
are connected over a network), such systems can be cheaply evaluated over
the Internet. This provides a potential solution to the long-standing
problem of evaluating NLG systems. The first instalment of GIVE (GIVE-1)
evaluated five NLG systems on the performance of 1143 users, making it the
largest ever NLG evaluation effort to date in terms of human users.

\begin{figure}[t]
\centering
%\includegraphics[width=1 \columnwidth]{give_world_no_expl}
\includegraphics[width=0.75\columnwidth]{give_world_2}
\caption{Map of an example GIVE world.}
  \label{fig:give-development-world}
\end{figure}

Planning plays a central role in the GIVE task. For instance, consider the
example GIVE world shown map in Figure~\ref{fig:give-development-world}. In
this world, the user's task is to pick up a trophy in the top left room.
The trophy is hidden in a safe behind a picture; to access it, the user
must push certain buttons in order to move the picture out of the way, open
the safe, and open doors. The user must navigate the world and perform
these actions in the 3D client; the NLG system must instruct the user on
how to do this. To simplify both the planning and the NLG task, the world
is discretised into a set of tiles of equal size. The user can turn by 90
degree steps in either direction, and can move from the centre of one tile
to the centre of the next tile, provided the path between two tiles is not
blocked. Figure~\ref{fig:give-planning} shows the encoding of some of the
available GIVE domain actions in PDDL syntax. In the example, the shortest
plan to solve the task consists of 108 action steps, with the first few
steps as follows:
%
\begin{enumerate}
\item $\mathsf{turn}\textsf{-}\mathsf{left}(\mathsf{north},
\mathsf{west})$,
\item $\mathsf{move}(\mathsf{pos\_5\_2}, \mathsf{pos\_4\_2}, \mathsf{west})$,
\item $\mathsf{manipulate}\textsf{-}\mathsf{alarm}\textsf{-}\mathsf{button}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{b1, pos\_5\_2})$,
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{west}, \mathsf{north})$.
\end{enumerate}

\begin{figure}
\centering
\begin{minipage}{0.9\textwidth}
{\small%
\begin{verbatim}
(:action move
   :parameters (?from - position  ?to - position  ?ori - orientation)
   :precondition 
       (and (player-position ?from)  (player-orientation ?ori)
            (adjacent ?from ?to ?ori)  (not (alarmed ?to)))
   :effect 
       (and (not (player-position ?from))  (player-position ?to)))

(:action turn-left
   :parameters (?ori - orientation  ?newOri - orientation)
   :precondition 
       (and (player-orientation ?ori)  (next-orientation ?ori ?newOri))
   :effect 
       (and (not (player-orientation ?ori))  (player-orientation ?newOri)))

(:action turn-right
   :parameters (?ori - orientation  ?newOri - orientation)
   :precondition 
       (and (player-orientation ?ori)  (next-orientation ?newOri ?ori))
   :effect 
       (and (not (player-orientation ?ori))  (player-orientation ?newOri)))

(:action manipulate-alarm-button-off-on
   :parameters (?b - button  ?pos - position  ?alarm - position)
   :precondition 
       (and (state ?b off)  (player-position ?pos)  (position ?b ?pos)
            (controls-alarm ?b ?alarm))
   :effect
       (and (not (state ?b off))  (not (alarmed ?alarm))  (state ?b on)))
\end{verbatim}}%
\end{minipage}
\caption{Simplified PDDL actions for the GIVE domain.}
\label{fig:give-planning}
\end{figure}

Our description of GIVE as a planning problem makes it very similar to the
classic Gridworld problem (see, e.g., \citealt{Tovey-Koenig:2000}
or the 1998 edition of the IPC\footnote{See
\texttt{ftp://ftp.cs.yale.edu/pub/mcdermott/aipscomp-results.html}.}),
which also involves route finding through a two-dimensional world map with
discrete positions. As in Gridworld, the domain also requires the
execution of certain object-manipulation actions (e.g., finding keys and
opening locks in Gridworld, or pushing the correct buttons to open doors
and the safe in GIVE). However, the worlds we consider in GIVE tend to be
much bigger than the Gridworld instances used in the 1998 planning
competition, with more complex room shapes and more object types in the
world.

To be successful in GIVE, an NLG system must be able to compute plans of
the form described above. At a minimum, a discourse planner will call a
domain planner in order to determine the content of the instructions that
should be presented to the user. This relatively loose integration of
NLG system and planner is the state of the art of the systems that
participated in GIVE-1. However, it is generally desirable to integrate the
planner and the generation system more closely than this. For instance,
consider an NLG system that wants to generate the instruction sequence
``walk to the centre of the room; turn right; now press the green button in
front of you''. Experiments with human instruction givers \citep{stoia:08}
show that this is a pattern that they use frequently: the instruction
follower is made to walk to a certain point in the world where the
instruction giver can then use a referring expression (``the green
button'') that is easy for the follower to interpret. An NLG system must
therefore integrate discourse planning and planning in the domain of the
world map closely. On the one hand, the structure of the discourse is
determined by the needs of the NLG system rather than the domain plan; on
the other hand, the discourse planner must be aware of the way in which the
instruction ``turn right'' is likely to change the visibility of objects.
Even if an NLG system doesn't implement the generation of such discourse as
planning, it must still solve a problem that subsumes the domain planning
problem. For these reasons, we consider the GIVE domain planning problem as
a natural part of a GIVE NLG system.
